---
title: "main"
author: "Zuofu Huang"
output: html_document
---

# Housekeeping

```{r message=FALSE}
library(caret)
library(tidyverse)
library(clv)
library(TraMineR)
library(ggplot2)
# library(Matrix)
# library(glmnet)
library(matrixcalc)

data <- read.csv("/Users/zuofuhuang/Desktop/reduceSeq/Transportation data/cleaned_data_noNA.csv")[,-c(1:3)]
colnames(data) <- c(format(seq(as.POSIXct("00:00:00", format = "%T"), 
                 as.POSIXct("23:59:00", format = "%T"), by = "1 min"), "%H:%M"))

data <- data %>%
  mutate_if(is.character, as.factor)

extract_time <- function(string){
  as.numeric(substring(string, 2, 3)) * 60 + as.numeric(substring(string, 5, 6)) + 1
}
```


```{r}
source("clust_assign_func.R")

seq_distances <- read.csv("/Users/zuofuhuang/Desktop/reduceSeq/Transportation data/sequences_dist_matrix_om.csv")[,-1]

hclust_real <- stats::hclust(as.dist(seq_distances), method = "complete")
fit_real <- cutree(hclust_real, k = 23)
table(fit_real)

# OLD code
# cluster_assignment <- create_clusters_Dunn(seq_distances)

# Current
cluster_assignment <- reorder_cluster_assignment(fit_real)
cluster_assignment[cluster_assignment > 14] <- 15

data_CA <- data
data_CA$CA <- as.factor(cluster_assignment)
```


# LASSO

```{r}
# # Ran 12/28/2024 with seed 256
# set.seed(256)
# lasso_caret <- train(as.factor(CA) ~ .,
#                     data = data_CA,
#                     method = "glmnet",
#                     family = "multinomial",
#                     trControl = trainControl(method = "cv", number = 10),
#                     tuneGrid = expand.grid(alpha = 1,
#                                           lambda  = seq(0, 0.1, length.out = 101)))
#                                        # lambda = 0.003))
# 
# lasso_results <- coef(lasso_caret$finalModel, lasso_caret$finalModel$lambdaOpt)
# 
# sum(predict(lasso_caret, type = "raw") == cluster_assignment) 
# mat1 <- table(predict(lasso_caret, type = "raw"), cluster_assignment) # top row is actual 
```

```{r}
lasso_caret <- readRDS("lasso_full_model_data_seed256")
lasso_results <- readRDS("lasso_results_data_seed256")
```

## LASSO coefficients

```{r}
positions_larger_than_threshold <- function(lasso_object, index, threshold){
  return(
    (lasso_object[[index]]@Dimnames[[1]])[lasso_object[[index]]@i[which(abs((lasso_object[[index]]@x)) > threshold)]]
  )
}

get_length <- function(lasso_object, index, threshold){
  return(
    length(positions_larger_than_threshold(lasso_object, index, threshold))
  )
}

get_times <- function(lasso_object, index, threshold){
  return(
    extract_time(positions_larger_than_threshold(lasso_object, index, threshold))
  )
}
```

```{r}
# For each level of the multinomial distribution, which variables were selected?
# sapply(1:15, get_length, lasso_object = lasso_results, threshold = 3)
# positions_left <- sort(unique(unlist(sapply(1:15, get_times, lasso_object = lasso_results, threshold = 0.01))))
```



## Use reduced number of predictors (through the cut-off threshold) to make the prediction

Reducing the number of predictors actually improves prediction. The regression coefficient does not fully represent how much that variable affects the prediction.

```{r}
# positions_left <- sort(unique(unlist(sapply(1:15, get_times, lasso_object = lasso_results, threshold = 2))))
# data_CA_reduced <- data_CA[,c(positions_left, 1441)]
#   
# set.seed(7003)
# lasso_step2 <- train(as.factor(CA) ~ .,
#                     data = data_CA_reduced,
#                     method = "glmnet",
#                     family = "multinomial",
#                     trControl = trainControl(method = "cv", number = 10),
#                     tuneGrid = expand.grid(alpha = 1, lambda = seq(0, 0.05, length.out = 101)))
# 
# # lasso_step2$bestTune
# lasso_results_Step2 <- coef(lasso_step2$finalModel, lasso_step2$finalModel$lambdaOpt)
# 
# sum(predict(lasso_step2, type = "raw") == cluster_assignment)
# # table(predict(lasso_step2, type = "raw"), cluster_assignment)
# 
# positions_left_step2 <- sort(unique(unlist(sapply(1:15, get_times, lasso_object = lasso_results_Step2, threshold = 0))))
# 
# # NA occasionally introduced for picking up intercept?
# positions_left_step2 <- na.omit(positions_left_step2)
# 
# length(positions_left_step2)
```



## Automate the process and create a plot

```{r}
# all_thresholds <- c(10^(-8:-2),seq(from = 0.1, to = 3, by = 0.05))
# 
# all_accuracy <- rep(NA, length(all_thresholds))
# all_lengths <- rep(NA, length(all_thresholds))
# 
# 
# for (i in 1:length(all_thresholds)){
#   positions_left <- sort(unique(unlist(sapply(1:15, get_times, lasso_object = lasso_results, threshold = all_thresholds[i]))))
#   data_CA_reduced <- data_CA[,c(positions_left, 1441)]
#   
#   set.seed(7003)
#   lasso_step2 <- train(as.factor(CA) ~ .,
#                        data = data_CA_reduced,
#                        method = "glmnet",
#                        family = "multinomial",
#                        trControl = trainControl(method = "cv", number = 10),
#                        tuneGrid = expand.grid(alpha = 1, lambda = seq(0, 0.05, length.out = 101)))
#   
#   all_accuracy[i] <- sum(predict(lasso_step2, type = "raw") == cluster_assignment)
# 
#   lasso_results_Step2 <- coef(lasso_step2$finalModel, lasso_step2$finalModel$lambdaOpt)
#   positions_left_step2 <- sort(unique(unlist(sapply(1:15, get_times, lasso_object = lasso_results_Step2, threshold = 0))))
#   
#   # NA occasionally introduced for picking up intercept?
#   positions_left_step2 <- na.omit(positions_left_step2)
#   
#   all_lengths[i] <- length(positions_left_step2)
# }
# 
# thresholdedLASSO_data_results_seed7003 <- as.data.frame(cbind(all_thresholds, all_accuracy, all_lengths))
# 
# colnames(thresholdedLASSO_data_results_seed7003) <- c("Threshold","Accuracy","Positions")

# saveRDS(thresholdedLASSO_data_results_seed7003,"thresholdedLASSO_data_results_seed7003")
```

```{r}
thresholdedLASSO_data_results <- readRDS("thresholdedLASSO_data_results_seed7003") %>%
  mutate(misclassify_percent = round(1-Accuracy/2892, 4) * 100)
```

```{r}
ggplot(thresholdedLASSO_data_results, aes(x = Threshold, y = 1-Accuracy/2892)) + 
  geom_point() + 
  geom_smooth(se = F) +
  ylab("Misclassification rate") +
  scale_x_continuous(limits = c(0,3), breaks = seq(0,3,0.25)) +
  scale_y_continuous(lim = c(0, 0.1), breaks = seq(0,0.1, 0.01)) + 
  theme_bw()

ggplot(thresholdedLASSO_data_results, aes(x = Threshold, y = Positions)) + 
  geom_point() + 
  geom_smooth(se = F) +
  ylab("Number of remaining positions") +
  scale_x_continuous(limits = c(0,3), breaks = seq(0,3,0.25)) +
  scale_y_reverse(lim = c(450,0)) + 
  theme_bw()

ggplot(thresholdedLASSO_data_results, aes(x = Positions, y = 1 - Accuracy/2892)) + 
  geom_point() + 
  geom_smooth(se = F) +
  scale_x_reverse(lim = c(450,0)) + 
  scale_y_continuous(lim = c(0, 0.1), breaks = seq(0,0.1, 0.01)) + 
  xlab("Number of remaining positions") +
  ylab("Misclassification rate") +
  theme_bw()
```


# Compare with random forests

```{r}
thresholdedLASSO_data_results <- readRDS("thresholdedLASSO_data_results_seed7003") %>%
  rename(positions = Positions, accuracy = Accuracy) %>%
  select(positions, accuracy) %>%
  mutate(method = "Thresholded LASSO")

rf_position_combined <- read.csv("../competitors/forests/rf_position_combined.csv")[,-1] %>%
  select(positions, accuracy) %>%
  mutate(method = "RF position combined")

rf_position_activity <- read.csv("../competitors/forests/rf_position_activity.csv")[,-1] %>%
  select(positions, accuracy) %>%
  mutate(method = "RF position activity")

combined <- rbind(thresholdedLASSO_data_results, rf_position_combined, rf_position_activity)

combined_reduced <- combined %>%
  filter(method != "RF position combined") %>%
  mutate(method = case_when(
    method == "Thresholded LASSO" ~ "Thresholded LASSO",
    method == "RF position activity" ~ "Random forest",
    .default = NA
  ))

ggplot(data = combined_reduced, aes(x = positions, y = 1 - accuracy/2892, color = method)) +
  geom_point() +
  geom_smooth(se = F) +
  xlab("Number of remaining positions") +
  ylab("Misclassification rate") + 
  scale_x_reverse(lim = c(450,0)) + 
  scale_y_continuous(lim = c(0, 0.1), breaks = seq(0,0.1, 0.01)) + 
  theme_bw()
  
```


