---
title: "Why does reduceSeq work the way it works?"
author: "Zuofu Huang"
output: html_document
---

Goal: Our reduceSeq method uses manual thresholding to get rid of smaller coefficients that LASSO doesn't shrink to 0 (but the accuracy is maximized). What if we make the lambda value larger than the cross validation identifies?

# Housekeeping

```{r message=FALSE}
library(caret)
library(tidyverse)
library(ggplot2)
library(clv)
library(TraMineR)
library(matrixcalc)
library(latex2exp)

data <- read.csv("/Users/zuofuhuang/Desktop/reduceSeq/Transportation data/cleaned_data_noNA.csv")[,-c(1:3)]
colnames(data) <- c(format(seq(as.POSIXct("00:00:00", format = "%T"), 
                 as.POSIXct("23:59:00", format = "%T"), by = "1 min"), "%H:%M"))

data <- data %>%
  mutate_if(is.character, as.factor)

extract_time <- function(string){
  as.numeric(substring(string, 2, 3)) * 60 + as.numeric(substring(string, 5, 6)) + 1
}
```

```{r message=FALSE, warning=FALSE}
# "baseline" sequences
palette <- c("#009292","#FF6DB6","#B6DBFF","#FFB6DB","#490092","#999999","#DB6D00","#004949","#B66DFF","#6DB6FF","#920000","#924900","#006DDB")

set.seed(32)
sample_sequences <- data[sample(1:1929, 100),]
seqiplot(seqdef(sample_sequences, cpal = palette), border = NA, cex.legend = 0.5, tlim = 1:300, ylab = NA, xtlab = NA, sortv = "from.start")

# Example: seqiplot(seqdef(data), border = NA, cex.legend = 0.5, idxs = sample(which(cluster_assignment == 1), min(50,sum(cluster_assignment == 1))), ylab = NA, xtlab = NA, cpal = palette)
```

Outline:
1. Create pairwise distances through optimal matching.
2. Hierarchical clustering with average distance
3. Group by Dunn index


```{r}
source("clust_assign_func.R")

seq_distances <- read.csv("/Users/zuofuhuang/Desktop/reduceSeq/Transportation data/sequences_dist_matrix_om.csv")[,-1]

hclust_real <- stats::hclust(as.dist(seq_distances), method = "complete")
fit_real <- cutree(hclust_real, k = 23)
# table(fit_real)

# cluster_assignment <- create_clusters_Dunn(seq_distances)
cluster_assignment <- reorder_cluster_assignment(fit_real)

# Consider cases where the groups are not consolidated down.
cluster_assignment[cluster_assignment > 14] <- 15
```

```{r}
data_CA <- data
data_CA$CA <- as.factor(cluster_assignment)
```


```{r}
lasso_results <- readRDS("lasso_results_data_seed256")
```

## Histogram - For each level of the multinomial distribution, which variables were selected?

```{r}
magnitude <- log10(abs(lasso_results[[1]]@x))
distributions <- cbind(rep(1, length(magnitude)), magnitude)

for (j in 2:15){
  magnitude <- log10(abs(lasso_results[[j]]@x))
  distributions <- rbind(distributions, cbind(rep(j, length(magnitude)), magnitude))
}

distributions <- as.data.frame(distributions)
colnames(distributions) <- c("Cluster", "magnitude")

distributions$Cluster <- paste("Cluster", distributions$Cluster)

distributions$Cluster <- factor(distributions$Cluster, levels = c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5", "Cluster 6", "Cluster 7", "Cluster 8", "Cluster 9", "Cluster 10", "Cluster 11", "Cluster 12","Cluster 13", "Cluster 14", "Cluster 15"))
```

```{r}
ggplot(distributions, aes(x = magnitude)) +
  geom_histogram(binwidth = 0.5, fill = "white", colour = "black") +
  facet_wrap(~Cluster, ncol = 5) +  
  xlab("Distribution of LASSO coefficient size for each binary model") + 
  ylim(0, 40) +
  scale_x_continuous(breaks = c(-15,-10,-5,0),
                     labels = c(TeX("$10^{-15}$"),TeX("$10^{-10}$"), TeX("$10^{-5}$"),1),
                     limits = c(-18,2)) +
  theme_bw()

ggplot(distributions, aes(x = magnitude)) +
  geom_histogram(binwidth = 0.2, fill = "white", colour = "black") +
  facet_wrap(~Cluster, ncol = 5) +  
  scale_x_continuous(breaks = c(-6,-4,-2,0),
                     labels = c(TeX("$10^{-6}$"),TeX("$10^{-4}$"), TeX("$10^{-2}$"),1),
                     limits = c(-6,1)) +
  xlab("Distribution of LASSO coefficient size for each binary model") + 
  ylim(0, 15) +
  theme_bw()
```
