---
title: "Group LASSO"
author: "Zuofu Huang"
output: html_document
---

# Housekeeping

```{r message=FALSE}
library(caret)
library(tidyverse)
library(clv)
library(TraMineR)
library(matrixcalc)
library(gglasso)
library(grplasso)
library(fastDummies)
```

```{r}
set.seed(7003)

data <- read.csv("/Users/zuofuhuang/Desktop/reduceSeq/Transportation data/cleaned_data_noNA.csv")[,-c(1:3)]
colnames(data) <- c(format(seq(as.POSIXct("00:00:00", format = "%T"), 
                 as.POSIXct("23:59:00", format = "%T"), by = "1 min"), "%H:%M"))

data <- data %>%
  mutate_if(is.character, as.factor)

extract_time2 <- function(string){
  as.numeric(substring(string, 1, 2)) * 60 + as.numeric(substring(string, 4, 5)) + 1
}
```

Outline:
1. Create pairwise distances through optimal matching.
2. Hierarchical clustering with average distance
3. Group by Dunn index


```{r}
source("../clust_assign_func.R")

seq_distances <- read.csv("/Users/zuofuhuang/Desktop/reduceSeq/Transportation data/sequences_dist_matrix_om.csv")[,-1]

hclust_real <- stats::hclust(as.dist(seq_distances), method = "complete")
fit_real <- cutree(hclust_real, k = 23)
# table(fit_real)

# cluster_assignment <- create_clusters_Dunn(seq_distances) 
cluster_assignment <- reorder_cluster_assignment(fit_real)

# Get rid of the really small "abnormalities" groups perhaps.
cluster_assignment[cluster_assignment > 14] <- 15
```

```{r}
data_wide <- fastDummies::dummy_cols(data, remove_selected_columns = T)

lasso_groups <- sapply(colnames(data_wide), extract_time2)
```


### Grplasso

```{r}
# data_CA <- data_wide
# data_CA$CA <- as.factor(cluster_assignment)

# fit <- grplasso(which(CA==1) ~ ., data = data_wide, model = LogReg(), lambda = 20, index = index)
```


### GGLASSO

gglasso only supports a two-factor categorical predictor.

We could, however, write a loop that loops over all possible lambdas, and perform a one-vs-rest binary logistic regression.

https://machinelearningmastery.com/one-vs-rest-and-one-vs-one-for-multi-class-classification/

```{r}
convert <- function(assignments, goal){
  pos1 <- which(assignments == goal)
  pos2 <- which(assignments != goal)
  assignments[pos1] <- 1
  assignments[pos2] <- -1
  return(assignments)
}
```

 
 
```{r}
# No need to cross validate, because there is only one lambda present
each_lasso <- function(data, cluster_assignment, lasso_group, lambda, current_cluster){ # a certain lambda
   grlasso <- gglasso(data, convert(cluster_assignment, current_cluster), lambda = lambda, group = lasso_group,
                            loss = "logit", intercept = F)
   coefs <- grlasso$beta
   nonzero_coefs <- coefs[coefs[,1] != 0,] # These are the coefficients
   positions <- unique(extract_time2(names(nonzero_coefs))) # Used these positions
   
   result <- predict(grlasso, newx = data, type = "link") # To predict the binary probability here
   return(list("results" = result, "beta" = coefs, "predictive_positions" = positions))
}

one_lambda_run <- function(data, cluster_assignment, lasso_group, lambda){ # all possible cluster assignments with the same lambda
  ngroup <- length(unique(cluster_assignment))
  
  result_matrix <- lapply(1:ngroup, each_lasso, data = data, cluster_assignment = cluster_assignment, lasso_group = lasso_groups, lambda = lambda)
  
  probs_matrix <- sapply(result_matrix, "[[", 1)
  beta_matrix <- sapply(result_matrix, "[[", 2)
  position_matrix <- sapply(result_matrix, "[[", 3)
  
  prediction <- apply(probs_matrix, MARGIN = 1, which.max)
  accuracy <- sum(prediction == cluster_assignment)/length(cluster_assignment)
  
  return(list("accuracy" = accuracy, "prediction" = prediction, "probs_matrix" = probs_matrix, "beta_matrix" = beta_matrix, "position_matrix" = position_matrix))
}

multinomial_lasso <- function(data, cluster_assignment, lasso_group, lambda_grid){
  multi_lambda_result <- lapply(lambda_grid, one_lambda_run, data = data, cluster_assignment = cluster_assignment, lasso_group = lasso_groups)
  return(multi_lambda_result)
}

get_multi_lasso <- function(multi_lambda_result, lambda_grid){
  accuracy <- sapply(multi_lambda_result, "[[", 1)
  best_lambda <- lambda_grid[which.max(accuracy)]
  
  positions_across_lambdas <- lapply(multi_lambda_result, "[[", 5)

  number_of_positions <- c()
  positions_list <- list()
  for (i in 1:length(positions_across_lambdas)){
    number_of_positions[i] <- length(unique(unlist(positions_across_lambdas[[i]])))
    positions_list[[i]] <- as.vector(unique(unlist(positions_across_lambdas[[i]])))
  }
  
  plot(number_of_positions, accuracy, type = "b")
  
  return(list(best_lambda, lambda_grid, accuracy, number_of_positions, positions_list))
}
```

**To ensure reproducibility, run all chunks before this at once. There is a seed 7003 at the top.**

```{r}
large_results <- multinomial_lasso(data = as.matrix(data_wide), cluster_assignment, lasso_groups, lambda_grid = 10^seq(from = -5, to = -1, length.out = 40))
res <- get_multi_lasso(large_results, 10^seq(from = -5, to = -1, length.out = 40))

# saveRDS(large_results, "groupLASSO_data_largeresults")
# saveRDS(res, "groupLASSO_data_res")
```

**To ensure reproducibility, run all chunks before this at once. There is a seed 7003 at the top.**

Rerun on Dec 28, 2024

```{r}
large_results <- readRDS("groupLASSO_data_largeresults")
res <- readRDS("groupLASSO_data_res")
```

```{r}
res_dat <- cbind(res[[2]], res[[3]], res[[4]])
colnames(res_dat) <- c("lambda","accuracy","position")
```


TLDR: Need about hundreds of positions collectively to get 98%. ~50 positions only get you ~88% accuracy.

```{r}
library(ggplot2)

ggplot(res_dat, aes(x = position, y = 1-accuracy)) +
  geom_point() +
  geom_smooth(se = F) +
  xlab("Number of remaining positions") +
  ylab("Misclassification rate") +
  scale_x_reverse(breaks = seq(1440,0, by = -240)) +
  scale_y_continuous(lim = c(0, 0.3), breaks = seq(0, 0.3, 0.05)) + 
  theme_bw()
```

